### 2014-10-22 tikz.tex
externalised Tikz node styling

%!TEX root=paper.tex

% basics

\tikzset{
  box/.style            = { rectangle, minimum width=2cm, minimum height=1cm },
  border/.style         = { thick, draw=black },
  round/.style          = { rounded corners=3mm },
}

% system

\tikzset{
  artifact/.style       = { box, font=\ttfamily },
  support/.style        = { box, font=\itshape },
  user/.style           = { artifact, fill=black!15!white },
  dom/.style            = { support,  fill=black!30!white },
  platform/.style       = { support,  fill=black!55!white, text=white },
  native/.style         = { support,  fill=black!70!white, text=white },
  generated/.style      = { artifact, border },
  resource/.style       = { artifact, border },
  representation/.style = { support, border, round }
}

% UML

\tikzset{
  class/.style={
    box,
    border,
    text centered,
    text width=1.75cm
  },
  relation/.style       = { thick, draw=black, <-},
  subclass/.style       = { relation, >=open triangle 60 },
  aggregate/.style      = { relation, >=open diamond },
  composite/.style      = { relation, >=diamond },
  dependency/.style     = { relation, >=latex, -> }
}