### 2014-11-13 paper.tex
another round of improvements

\documentclass[conference]{IEEEtran}
\usepackage[pdftex]{graphicx}
\usepackage[cmex10]{amsmath}
\interdisplaylinepenalty=2500
\usepackage{array}
\usepackage{mdwmath}
\usepackage{mdwtab}
\usepackage{eqparbox}
\usepackage[font=footnotesize]{subfig}
\usepackage{fixltx2e}
\usepackage{url}
\hyphenation{}

\input{packages}

\begin{document}

\title{
\NAME: A framework for \\
Efficient Intrusion Detection\\
in the Internet of Things
}

\author{\IEEEauthorblockN{Christophe Van Ginneken, Jef Maerien, Danny Hughes, 
Christophe Huygens, Wouter Joosen}%
\IEEEauthorblockA{iMinds-DistriNet, KU Leuven\\
3001 Leuven, Belgium\\
\{firstname.lastname\}@cs.kuleuven.be}}

\maketitle

\begin{abstract}

When intrusion prevention fails, intrusion detection serves as a second layer
of defense. It looks for patterns \& anomalies that can indicate malicious
behavior. But supporting an adequate set of intrusion detection algorithms
imposes significant overhead on resource-constraint devices such as those in
the Internet of Things. Fusion of the algorithms, to optimize sharing of
resources, can alleviate this, yet, proves to be time-consuming, repetitive and
error-prone. To address this problem \NAME provides a framework for the
development of efficient intrusion detection systems. It consists of a domain
specific language, to formally describe intrusion detection algorithms on a
functional level, and a code generator to produce more optimally organized
source code. A side-by-side comparison shows that generated code is smaller and
reduces message passing overhead, execution time and memory footprint in
comparison to a straightforward sequential implementation of the same
algorithms.

\end{abstract}

\section{Introduction}

% context

% IoT + threat

The internet of things (IoT) holds a great potential to positively influence
our daily work and life through domotics, assisted living, e-health and
enhanced learning \cite{atzori2010internet}. But with this potential the IoT
also presents a significant threat: by opening our smart homes and personal
data to all these interconnected devices, we also open them to everyone who is
able to break the virtual locks that protect them (i.e.\ passwords, encryption
keys,\dots). IoT devices typically have limited batteries, processing power and
memory. These properties make it hard to add any additional service, including
security.

% classic networks: fw and central ids

In classic networks, firewalls focus on the outer perimeter of the network,
filtering unwanted packets and protecting the entire internal network. But
attacks on flaws in services can pass unnoticed. This is where intrusion
detection (ID) comes into play: an intrusion detection systems (IDS) monitors
all traffic that passes through the firewall, looking for patterns of malicious
activity, and optionally, after detecting such pattern, alerting the firewall,
allowing it to take corrective actions \cite{denning1987intrusion}.

% local ids -> localize -> limited resources

In wireless networks of resource-constraint devices, which make up a
significant part of the IoT, it is not possible to have this single point in
the network that can oversee all traffic \cite{mishra2004intrusion}. Every
device has to implement its own lines of defense. But making the IDS a local
service on each device requires local resources, which are a scarce commodity
for IoT devices.

Attackers have access to a large set of possible attack vectors
\cite{aschenbruck2012security}, ranging from the wireless network, over the
medium access control and routing, up to the device itself. Each level presents
different opportunities to manipulate data, eavesdrop, perform a form of denial
of service or even access the network in a legitimate way. For each of these
attacks an algorithm needs to look for patterns or anomalies. An IDS for IoT
devices therefore requires a large number of implemented algorithms for an
adequate coverage.

% gap analysis

Current developments in ID focus on programming frameworks to structure the
implementation of algorithms \cite{valero2012di} and offer the required basic
functional components to implement them \cite{krontiris2008lidea}. The first
aspect that is missing, is that they don't offer a way to optimize the usage of
resources. A solution that supports the integration of multiple algorithms on
IoT devices needs to avoid accumulating their impact on the available resources.

A second omission is support for systematic reuse to create different
configurations for different devices, both in hard- and software. Addressing
this in a transparent and automated way is important for any kind of
development in this area.

% scientific contribution

% 1. framework: language + code generator

The first contribution of this paper is a framework called \NAME. It enables
the formal description of ID algorithms on a functional level and provides a
code generator to produces corresponding source code for a given platform and
configuration. This way, \NAME addresses the heterogeneous nature of both the
IoT and ID algorithms.

% 2. FCF to optimise

The second contribution introduces functional code fusion (FCF) as a source
code generation paradigm to solve the problem of hard to reuse and combine ID
algorithms. FCF identifies common data and functions, and organizes code to
eliminate redundant iterations, tests and computations. Results show improved
image size, usage of the wireless radio, execution time and memory usage.

% structure

The remainder of this paper proceeds as follows: section \ref{classification}
analyses and classifies ID algorithms. Section \ref{pattern} looks for patterns
in the implementation of these classes. Section \ref{problems} identifies the
problems arising from these patterns. Section \ref{design} describes the design
we applied to construct \NAME, its domain specific language (DSL) and code
generator. Section \ref{evaluation} discusses and evaluates an implementation.
Section \ref{related} explores related work in the field of DSLs and code
generation. Finally, section \ref{conclusion} summarizes our findings, draws
conclusions and identifies topics for future work.

\section{A Classification of ID Algorithms}
\label{classification}

Two aspects, currently missing from ID frameworks, that \NAME wants to address,
are optimization of resources and support for systematic reuse. This section
uses a classification to define the scope of \NAME, detailing which types of
algorithms are good candidates for the proposed approach.

Research into ID in wireless networks typically focuses on two major topics:
the devices as single entities and the network as a group of such devices. Both
are important: a group cannot make a decision without members that detect
malicious behavior and a group-based decision is often needed because devices
in the network can miss certain events that could reveal intrusions. In a
wireless network, where not all participants are within each others' radio
range, wireless devices typically only have a partial view on the entire
network. This prohibits them to observe potentially crucial events in distant
parts of the network. The fact that resource-constraint devices typically apply
duty cycling to lower their energy consumption, further increases this problem,
introducing periods of time where they even don't observe local events.

This leads to two complementary classes of ID activities: local intrusion
detection and cooperative decision making.

\subsection{Local Intrusion Detection}
\label{detection}

There are different ways to construct a taxonomy for intrusion detection, but
common themes do appear. According to \cite{mishra2004intrusion} and
\cite{ioannis2007towards} there are three major categories: anomaly detection,
signature or misuse detection, and specification-based detection. In
\cite{alrajeh2013intrusion} the authors agree with this typology and add the
notion of hybrid intrusion detection systems and cross layer intrusion
detection systems as recent advances in research. Although these additional
categories seem promising, the authors admit that the impact on
resource-constrained devices will be too much. We therefore focus on the three
categories agreed upon by most sources.

\begin{LaTeXdescription}

  \item[Anomaly detection] considers a baseline profile and evaluates all
  activity according to this profile. It tries to flag aberrant behavior as a
  possible intrusion. The baseline profile can evolve over time, especially in
  wireless and mobile environments. Single events should therefore not result
  in absolute decisions.

  \item[Signature or misuse detection] starts with known patterns of attacks.
  It tries to find evidence for intrusions in the actions of and communication
  with a device. Because attacks often use valid ways to communicate with
  services on the devices, this approach can produce false positives.

  \item[Specification-based detection] uses constraints that describe the
  correct behavior of the system. Although that this approach can detect
  previously unknown attacks, it can sometimes not detect specific attacks that
  anomaly or misuse detection can.

\end{LaTeXdescription}

\subsection{Cooperative Decision Making}
\label{coorperative}

These taxonomies focus on the act of detecting intrusions themselves. In a
distributed environment, communication between participants may be even more
important.

Because not all devices are always actively participating in the network, they
can miss events that would otherwise lead them to detect intrusions. It is
hardly impossible for a single device to detect a complex attack by itself.
Therefore a second important research topic consists of cooperative algorithms
to combine information from single identities into a group-based decision about
alleged intrusions. We identify two ways to construct cooperative decision
making and focus on the way participants exchange information: broadcast and
interactive communication.

\begin{LaTeXdescription}

  \item[Broadcast communication] can allow implementing a distributed,
  cooperative algorithm locally. The reputation-based detection algorithm
  introduced in \cite{ganeriwal2008reputation} illustrates this. It uses the
  same principle of a watchdog also found in \cite{mishra2004intrusion}. Using
  this algorithm, devices can exchange information about the reputation of
  other devices, solely using one-way broadcast messages. Combining this
  information allows them to decide about their trust in a given device, based
  on more than their own, often partial, observations.

  \item[Interactive communication] allows algorithms to exchange information to
  reach a consensus. In \cite{krontiris2009cooperative}, the authors first
  present a theoretical foundation to analyze cooperative algorithms. Given
  these foundations they continue to present an algorithm consisting of five
  phases, which essentially implement a voting system based on the Guy Fawkes
  protocol \cite{anderson1998new}, to identify an intruder. It enables the
  exchange of suspected intruder information and allows for distributed
  authentication of those \emph{votes}. Based on these authenticated votes, the
  participants can make a distributed decision about the jointly identified
  intruder.

\end{LaTeXdescription}

\subsection{Properties and Scope of \NAME}
\label{scope}

Algorithms that fit this classification are typically platform independent.
They analyze network messages, looking for patterns that can reveal intrusions
and all deal with the same information: other devices in the network. To be
able to formally describe ID algorithms and target the heterogeneous IoT, \NAME
requires ID algorithms to expose these properties.

A class of ID algorithms that does not have these properties is software
attestation \cite{seshadri2008sake}. It enables exchanging information about
the software that is running on a device and identifying devices with modified
content. These algorithms are highly platform dependent and have no data in
common with other algorithms, as they look for the effects of an intrusion, not
the intruder itself.

Figure \ref{fig:classification} visually summarizes the dual classification and
scope of \NAME.

\begin{figure}[ht]
  \centering
  \scalebox{.85}{
\begin{tikzpicture}[
  sibling distance=20mm,node distance=40mm
  ]
  \node (detection)             [align=center] {Detecting\\Intrusions}
    child {node (anomaly)       [align=center] {Anomaly}}
    child {node (signature)     [align=center] {Signature\\Misuse}}
    child {node (specification) [align=center] {Specification}};
  \node (decision)              [align=center,right=of detection] {Decision\\Making}
    child {node (broadcasting)  [align=center] {Broadcasting}}
    child {node (interactive)   [align=center] {Interactive}};
\end{tikzpicture}
}
  \caption{Classification of ID Algorithms}
  \label{fig:classification}
\end{figure}

\section{A Design Pattern for ID Algorithms}
\label{pattern}

Given this classification, we now introduce a design pattern that identifies
common functional aspects of these algorithms. It will allow us to define a
strategy for optimizing the combination of these algorithms. From the
classification and a survey of algorithms, we extract two activities common to
most algorithms: the inspection of network traffic and the evaluation of
thresholds.

Network traffic is the carrier along which malicious payloads reach a device
directly. Indirectly a device can also detect aberrant patterns of
communication by other devices, as a result of a successful attack.

Not all algorithms can decide if an attack is real or not, based on a single
event. Signature and misuse detection can in some cases, but anomaly detection
typically deals with some level of uncertainty. Also, in distributed situations
it is impossible to fully trust any neighboring device. In these situations one
has to rely on statistics and thresholds to decide when to drop trust in a
given device.

But simply checking a threshold after updating some counter, is not a good
strategy. A simple example is that of a watchdog \cite{mishra2004intrusion},
observing the actions of other devices. There are valid reasons why a device
can miss a theoretical window in which one expects activity, i.e.\ the wireless
network being the most predominantly unreliable component. Directly reacting on
a basic threshold can result in false positives. A way to overcome this is to
decouple the detection of problems and the decision what to do about them.

This decoupling of detection and decision is the core of the design pattern for
ID algorithms. Listing \ref{alg:id-algo-pattern} presents the design pattern in
pseudo-code.

\begin{figure}[ht]
\captionof{lstlisting}{ID Algorithm Pattern\label{alg:id-algo-pattern}}
\begin{algorithmic}[1]
  \Require{devices: global storage for device information}
  \Function{process\_message}{$msg$}
    \ForEach{$byte \in msg$} \label{alg:id-algo-pattern-loop1}
     \State \dots \Comment{analyze byte-sequences}
    \EndFor
    \Let{$devices_x$}{value}  \Comment{optionally update device info}
    \State \Call{send}{$device_y$, ``info''} \Comment{optionally exchange info} \label{alg:id-algo-pattern-send1}
  \EndFunction
  \State
  \Function{do\_housekeeping}{}
    \ForEach{$device \in devices$} \label{alg:id-algo-pattern-loop2} \label{alg:id-algo-pattern-common-data}
      \If{$device > \dots$} \Comment{validate recorded value}
        \State \dots \Comment{take actions}
        \State \Call{send}{$device$, ``info''} \Comment{communicate} \label{alg:id-algo-pattern-send2}
      \EndIf
    \EndFor
  \EndFunction
  \State
\end{algorithmic}
\end{figure}

Applying the pattern in the context of multiple algorithms, extends its scope
and a pattern such as described in listing \ref{alg:id-algo-application} will
typically appear.

\begin{figure}[ht]
\captionof{lstlisting}{Application pattern of multiple ID algorithms\label{alg:id-algo-application}}
\begin{algorithmic}[1]
  \Let{$msg$}{$network$} \Comment{all observed messages}
  \ForEach{$algorithm \in algorithms$}
    \State \Call{algorithm.progress\_message}{msg}
  \EndFor
  \State \dots
  \Comment{at a given interval}
  \ForEach{$algorithm \in algorithms$}
    \State \Call{algorithm.do\_housekeeping}{}
  \EndFor
\end{algorithmic}
\end{figure}

The pattern calls the different, linked implementations in sequence in two
situations: first, when the device observes new messages on the network and,
secondly, at a regular interval to allow for internal housekeeping, such as
threshold evaluation.

\section{Inherent Problems}
\label{problems}

These patterns can now help in identifying the problem areas that tend to
result in suboptimal code when combining multiple ID algorithms. Besides these
code-level problems, operational problems also hinder the implementation of an
IDS for IoT devices.

Two issues arise from the nesting of loops that originate from the combination
of the ID algorithm pattern and the application pattern: loops and common data.
A third issue is the use of the network.

\begin{LaTeXdescription}

  \item[Loops] by themselves are not a problem, but different loops, that
  perform the same operations in sequence, can incur an overhead. An example
  from the ID algorithm pattern is the handling of messages: each algorithm
  parses each observed message on the network, processing the same byte stream,
  performing comparable operations to find similar patterns.

  \item[Common data] cross-cuts the different algorithms, that have to store
  information about devices within its own scope. This results in a lot of
  duplicated information in memory.

  \item[The network] allows algorithms to operate on a distributed scale. But
  scattered calls to the \ttt{SEND} function cause multiple accesses to the
  wireless radio, keeping it active and energy consuming more than strictly
  needed.

\end{LaTeXdescription}

Two operational aspects are orthogonal to these problems and amplify them:

\begin{LaTeXdescription}

  \item[Reimplementing] the algorithms manually to eliminate these issues is
  not a one-time cost. For example due to the release of a new version of an
  algorithm, or simply a reselection of algorithms due to a changing risk
  analysis. Also, changing, or simply implementing the algorithms from scratch,
  holds the risk of making mistakes.

  \item[The heterogeneity of IoT devices] results in different software stacks,
  each with for example their own network API\@. This also explains why there
  are no implementations readily available to developers: there is no single
  platform to target for researchers.

\end{LaTeXdescription}

\section{Design}
\label{design}

The previous sections identified loops, common data and the network as three
important code-level problems. These problems are inherent to ID algorithms.
But at the same time they conflict with their target environment, the IoT\@.
\NAME provides a framework to efficiently implement an IDS for the IoT and
offers a solution for the identified problems. It consists of a DSL, source
code generator and software library. Figure \ref{fig:design} visualises the
overall design.

\begin{figure}[ht]
  \centering
  \scalebox{.85}{
\begin{tikzpicture}[
  node distance=7mm,
  ,>=stealth,thick,black!50,text=black,
  every new ->/.style={shorten >=1pt}
]
  \node (dsl)  [resource]                                          {DSL};
  \node (cg)   [representation,right=of dsl,align=center]          {Code\\Generator};
  \node (code) [resource,right=of cg]                              {Code};
  \node (lib)  [resource,below=of code,align=center,yshift=0.5cm]  {Software\\Library};

  \draw [->] (dsl) -> (cg)   node [midway,above=25pt]              {intent};
  \draw [->] (cg)  -> (code) node [midway,above=15pt,align=center] {platform\\ + language};

  \draw [rounded corners,->]
        ($ (lib.west) $)
        -- ++(-0.5cm, 0cm)
        -- ++( 0cm,   1cm)
        -- ($ (code.west) + (0cm,-0.22cm) $);

\end{tikzpicture}
}
\caption{Overall Design}
\label{fig:design}
\end{figure}

The DSL enables researchers to formally describe ID algorithms, without
focusing on any specific language or platform. A code generator can avoid the
identified problems concerning loops, common data and network access, as the
formal description provides the underlying intent, not the technical
implementation. It can also target different languages and platforms, while
reusing a software library that provides services including parsing, network
access wrapping,\dots

A DSL and code generator defines a development strategy. This implies choices
with respect to flexibility, update cost and development effort. We will first
position \NAME and motivate its design with respect to these trade offs.
Second, we will look at the principal design components of the DSL and how it
deals with loops and common data. Third, we discuss the structure of the code
generator and introduce its software library, that offers a way to deal with
the network access issue.

\subsection{\NAME as a Development Strategy}
\label{positioning}

Figure \ref{fig:positioning} shows a classification of possible development
strategies for projects involving wireless devices.

\newcommand{\ball}[5][above,yshift=+5pt]{
\tdplottransformmainscreen{#2}{#3}{#4}
\shadedraw[tdplot_screen_coords, ball color=lightgray,draw=lightgray,shading angle=-90]
  (\tdplotresx,\tdplotresy) circle (0.20)
  node[text=gray,font=\footnotesize,#1]{#5};
}

\begin{figure}[hb]
  \centering
  \scalebox{.85}{
\tdplotsetmaincoords{110}{-10}
\begin{tikzpicture}[scale=1,tdplot_main_coords]
  % axis
  \draw[thick,->] (0,0,0) -- (7,0,0) node[anchor=north east,xshift=10mm]{$flexibility$};
  \draw[thick,->] (0,0,0) -- (0,7,0) node[anchor=north west]{$update\ cost$};
  \draw[thick,->] (0,0,0) -- (0,0,4) node[anchor=south]{$development$};
  % points
  \coordinate (O)        at (0,0,0);
  \coordinate (ideal)    at (6,0,0);
  \coordinate (vm)       at (2,2,2);
  \coordinate (modular)  at (4,4,4);
  \coordinate (reconfig) at (4,2,4);
  \coordinate (full)     at (6,6,4);
  \coordinate (cg)       at (6,4,2.5);
  % "lines"
  \draw[dashed] (2,0,0)    -- (2,2,0) -- (0,2,0);
  \draw[dashed] (4,0,0)    -- (4,4,0) -- (0,4,0);
  \draw[dashed] (6,0,0)    -- (6,6,0) -- (0,6,0);
  \draw[dashed] (vm)       -- (2,2,0);
  \draw[dashed] (modular)  -- (4,4,0);
  \draw[dashed] (full)     -- (6,6,0);
  \draw[dashed] (reconfig) -- (4,2,0);
  \draw[dashed] (cg)       -- (6,4,0);
  % "balls"
  \ball{6}{0}{0}{ideal}
  \ball{2}{2}{2}{virtual machine}
  \ball[left,xshift=-5pt]{4}{4}{4}{\shortstack{modular\\native image}}
  \ball{4}{2}{4}{\shortstack{reconfigurable\\virtual machine}}
  \ball[right,xshift=+5pt,yshift=+5pt]{6}{6}{4}{full native image}
  %\ball[right,xshift=+5pt,text=black]{6}{4.5}{3}{code generation};
  % 5. cg
  \tdplottransformmainscreen{6}{4}{2.5}
  \shadedraw[tdplot_screen_coords, ball color=gray,draw=black,shading angle=-90]
    (\tdplotresx,\tdplotresy) circle (0.20)
    node[right,xshift=+5pt,text=black,align=center,font=\footnotesize]{code generation\\+ functional code fusion};
  % "box around our solution" ;-)
  \tdplottransformmainscreen{0}{0}{0}
  \draw[tdplot_screen_coords] (4.95,1) rectangle (8.5,0.25);

\end{tikzpicture}
}
\caption{Classification of Development Strategies for Wireless Devices (based on \cite{sadilek2007energy})}
\label{fig:positioning}
\end{figure}

Three properties can classify development strategies \cite{sadilek2007energy}:
flexibility, update cost and development effort.

\begin{LaTeXdescription}

  \item[Flexibility] is a measure for the number of applications one can deploy
  on a device without physical access.

  \item[Update cost] is equivalent to the energy required to deploy a new
  version of an application on a device.

  \item[Development effort] covers the time to write, test and deploy an
  application.

\end{LaTeXdescription}

The \emph{ideal} situation is that were one ends up with a maximal flexibility
and no update nor development cost. \NAME starts from the full native image,
allowing full flexibility, and aims to reduce both the update and development
cost: first, by applying functional code fusion, \NAME optimizes the code
organization, which has a positive effect on the code size and hence also on
the update cost. Second, by enabling reuse on a functional level and automating
the writing phase, the development cost is also lowered.

\subsection{\NAME as a Domain Specific Language}
\label{dsl-design}

It's important to remember that a DSL is merely a \emph{thin veneer over a
(semantic) model} \cite{fowler2010domain}. Its sole purpose should be
populating the model. \NAME's DSL and model focuses on avoiding constructions
that result in problems concerning loops, common data and the use of the
network. To do this, \NAME uses an event-based approach to defining
functionality.

\subsubsection{Loops}

Loops are a major contributor to the problem of algorithm fusibility. Loops in
program code are technical constructs and in fact often hide the intended
functionality.

The example of parsing network messages illustrates this. Each algorithm deals
with parsing in the same way: loop over the bytes in the payload and look for a
pattern that identifies what the algorithm is looking for. This is in fact a
technical implementation choice. The underlying functional intent is to react
when such a pattern is present in a message. The loop has \emph{activated} this
functionality, but the real goal is to \emph{passively} react on an event, not
\emph{actively} look for it. Event handling therefore can transform
constructions with loops to their functional meaning, allowing identification
and fusion of common functionality.

The same goes for actively polling some variable until it reaches a certain
threshold. When transformed to a reaction to an event, it eliminates the need
for a technical loop-construction.

\subsubsection{The Nodes Domain}

All algorithms deal with information about other devices in the network. To
centralize this information \NAME provides a unified \ttt{nodes} concept as
part of its DSL and exposes it as an object to the implementer.

Algorithms can extend nodes with their own properties. This way, only one
instance for every known device will be present in memory. Finally, nodes can
define a scope for a function to execute in. This way, nodes also offer a way
to i.e.\ iterate all known devices, again without the need for a loop.

\subsubsection{Unified Messaging}
\label{dsl-unified-msg}

Access to the network, scattered around an algorithm, causes the wireless radio
to be more active than needed. Delaying the sending of messages and grouping
them, can lower this overhead.

Sending and receiving of network messages is tightly integrated with the
previously mentioned concepts of event handling and the Nodes domain. An
underlying software library can handle all interaction with the network,
triggering functions by means of events when parsing new messages and exposing
an API to send messages through the Nodes domain. This approach gives the
framework complete control over communication, allowing optimized marshaling,
grouping,\dots

\subsubsection{An Example}

Listing \ref{lst:watchdog} illustrates some of the features mentioned above,
using a small but representative example of an algorithm: a
watchdog \cite{mishra2004intrusion}. Like a heartbeat, nodes can broadcast a
message at a regular interval, allowing other nodes to monitor its
availability. Missing heartbeats could indicate problems, such as physical
capturing of a device.

\lstinputlisting[
  language=foo-lang,
  float,
  basicstyle=\footnotesize\ttfamily\color{black},
  label=lst:watchdog,
  caption=Watchdog in \NAME
]{src/watchdog.foo}

\subsection{\NAME as a Code Generator}
\label{code-generator-design}

To convert the DSL-based ID algorithms into production-ready code, we apply
code generation techniques. Although different approaches here are applicable,
like template-based text expansion, XSLT transformations or CodeDom
\cite{dollard2004code}, we believe that the functional transformations require
a richer model.

A semantic model \cite{fowler2010domain}, tied closely to the DSL itself, can
fulfill this role. Semantic transformations evolve the model into a more
syntax-oriented model, comparable to the idea of CodeDom. From there on,
syntactic transformations can further evolve the model into source code. Figure
\ref{fig:code-generation} shows the process that \NAME implements.

\begin{figure}[ht]
  \centering
  \scalebox{.85}{
\begin{tikzpicture}[
  node distance=7mm,
  ,>=stealth,thick,black!50,text=black,
  every new ->/.style={shorten >=1pt}
]
  \node (dsl)  [resource]                                 {DSL};
  \node (sm)   [representation,right=of dsl,align=center] {Semantic\\Model};
  \node (cm)   [representation,right=of sm,align=center]  {Syntactic\\Model};
  \node (code) [resource,right=of cm]        {Code};

  \draw [->] (dsl) -> (sm)   node [midway,above=15pt]              {parse};
  \draw [->] (sm)  -> (cm)   node [midway,below=10pt,align=center] {semantic\\transform}
                             node [midway,above=15pt,align=center] {FCF};
  \draw [->] (cm)  -> (code) node [midway,above=15pt]              {emit code};

  \draw [rounded corners,->]
        ($ (sm.east) - (5mm,5mm) $)
        -- ++(0,-.5)
        -| ($ (sm.west) + (5mm,-5mm) $)
           node [near start, below=5pt] {type inference};

  \draw [rounded corners,->]
        ($ (cm.east) - (5mm,5mm) $)
        -- ++(0,-.5)
        -| ($ (cm.west) + (5mm,-5mm) $)
           node [near start, below=1pt,align=center] {syntactic\\transform};

\end{tikzpicture}
}
\caption{Overview of \NAME's code generation process}
\label{fig:code-generation}
\end{figure}

\subsubsection{Semantic Model}

The semantic model is the core of the solution. It offers a way to describe ID
algorithms in a way that allows semantic reinterpretation through functional
code fusion. This way, reorganizing the functionality of different algorithms
allows for more optimal execution and memory usage. Hence the name \NAME of the
framework: intrusion detection functionality organization optimization.

Figure \ref{fig:meta-model} shows the central part of the semantic model as
implemented for \NAME.

\begin{figure}[ht]
  \centering
  \scalebox{.85}{
\begin{tikzpicture}[
  edge from parent fork down,
  sibling distance=5mm, level distance=20mm, node distance=6mm
  ]
  \node (model) [class,fill=black!10!white] {Model}
      child {node (module) [class,fill=black!10!white] {Module} edge from parent [composite]
        child {node (strategy) [class,align=center,xshift=-10.5mm,fill=black!10!white] {Execution\\Strategy} edge from parent [composite]
          child {node(every) [class,xshift=-9mm,fill=black!10!white] {Every} edge from parent [subclass]}
          child {node (when) [class,xshift=9mm,fill=black!10!white] {When} edge from parent [subclass]}
        }
        child {node (function) [class,xshift=10.5mm] {Function} edge from parent [composite]}
      };
  \node (constant)  [class,right=of module]   {Constant};
  \node (scope)     [class,left=of strategy]  {Scope};
  \node (statement) [class,right=of function] {Statement};
  \node (nodes)     [class,left=of module,xshift=-13mm]    {Nodes\\Domain};

  \draw [aggregate]  (module)   -- (nodes);
  \draw [composite]  (module)   -- (constant);
  \draw [dependency] (strategy) -- (scope);
  \draw [composite]  (function) -- (statement);

  \draw [composite]  (nodes)    -- (scope);
  \draw [dependency] (strategy) -- (function);
\end{tikzpicture}
}
\caption{Partial Semantic Model for \NAME. Shaded classes indicate the \emph{event/scheduling} construction around functions to allow interpretation.}
\label{fig:meta-model}
\end{figure}

The concept driving this model is the structure starting at the \emph{Model} up
to the \emph{Execution Strategy}. This structure encompasses the
\emph{Function}s and underlying \emph{Statement}s and allows for interpretation
and fusion of the algorithms.

\subsubsection{Functional Code Fusion (FCF)}

Thanks to the functional description of events and the effects triggered by
them, the code generator is able to identify the same actions across different
algorithms. This way it is possible to extract the parsing of incoming
messages, group actions on the same sets of data or scheduled function
executions.

FCF typically looks for patterns in the semantic model and performs model
transformations to change these patterns to a more optimized organization. It
first performs semantic transformations on the semantic model itself and ends
then transforms the semantic model into a syntactic model, encoding the
semantics into technical syntax constructs.

One example of a semantic transformation is type inference, which supports the
platform independence of the framework.

\subsubsection{Syntactic Model}

The syntactic model is an AST for a virtual, expressive language. It contains
constructs and ideas found in different languages and programming paradigms. No
single existing programming language supports all features as is. Syntactic
transformations therefore morph the unsupported constructs into comparable
constructs supported by the targeted platform and language.

Depending on the target platform, some constructions aren't transformed into
actually corresponding code. In selected cases, the generator uses a software
library. For example when it needs to parse incoming messages or schedule
tasks. During the generation process, it looks for patterns that indicate any
of these actions and transforms them into calls to the \NAME software library.

\subsection{\NAME as a Software Library}
\label{software-lib-design}

A software library allows offloading of common implementation aspects, thus
avoiding small discrepancies in code and suboptimal use of shared resources.
Section \ref{dsl-unified-msg} illustrated this approach with a case for unified
messaging: parsing of incoming messages and grouping of outgoing messages.

Other common functionality also isn't generated as-is. The execution of
functions based on a schedule or the implementation of higher level conceptual
functions, such as \ttt{now} and \ttt{sha1}, are calls into the standard
software library that supports the code generator.

Optimization of such algorithms is not within the scope of \NAME. It also gives
freedom of choice to the users of the framework to select appropriate
implementations for this standard functionality.

\subsection{\NAME as a Framework}

\NAME reduces inefficiencies in network access, execution and memory usage,
that occur due to a selection of ID algorithms. To achieve this, it avoids
constructions that are hard to identify and reorganize, such as loops and
common data. A DSL, focussing on event-handling, enforces these restrictions
and enables formally describing ID algorithms.

The DSL is also platform and language independent. This allows the source code
generator to target different configurations without any change to the formal
descriptions of the algorithms.

A supporting software library allows separation of standard functionality that
is out of scope for the generator. Optimization of the implementation of this
library for the specific platform and target language is possible.

\section{Implementation and Evaluation}
\label{evaluation}

% table goes here, because a spanning table is placed on the next page, relative
% to where it's defined.
\begin{table*}[b]
  \centering
  \begin{tabular}{llrrrrrrr}
\toprule
              &            & base  & \multicolumn{2}{c}{watchdog} & \multicolumn{2}{c}{reputation} & \multicolumn{2}{c}{both} \\
\midrule
image
size (bytes)  & manual     & 10466 & 15530 &  +48\% & 13306 &  +27\% & 18334 &  +75\%\\
              & generated  & 10466 & 18352 &  +75\% & 16376 &  +56\% & 20998 & +101\%\\
\cmidrule{2-9}
              & difference &     0 &  2822 &  +18\% &  3070 &  +23\% &  2664 &  +15\%\\
\midrule
\multirow{2}{*}{\pbox{\textwidth}{total packets sent\\in a window of 90s}}
              & manual     &    20 &    51 & +155\% &    32 &  +60\% &    63 & +215\%\\
              & generated  &    20 &    49 & +145\% &    32 &  +60\% &    55 & +175\%\\
\cmidrule{2-9}
              & difference &     0 &    -2 &   -4\% &     0 &   0\% &    -8 &   -13\%\\
\midrule
\multirow{2}{*}{\pbox{\textwidth}{total bytes sent\\in a window of 90s}}
              & manual     &   476 &  1933 & +306\% &   860 &  +81\% &  2317 & +387\%\\
              & generated  &   476 &  1897 & +299\% &   884 &  +86\% &  2161 & +354\%\\
\cmidrule{2-9}
              & difference &     0 &   -36 &   -2\% &    24 &   +3\% &  -156 &   -7\%\\
\midrule
\multirow{2}{*}{\pbox{\textwidth}{event loop\\average duration ($\mu$s)}}
              & manual     &    48 &    94 &  +96\% &    88 &  +83\% &   149 & +210\%\\
              & generated  &    48 &   121 & +152\% &   121 & +152\% &   138 & +188\%\\
\cmidrule{2-9}
              & difference &     0 &    27 &  +29\% &    33 &  +38\% &   -11 &   -7\%\\
\bottomrule
  \end{tabular}
  \caption{Results and Side-by-Side Comparison.}
  \label{tbl:results}
\end{table*}

This section discusses a prototype implementation and evaluates its
performance. We first present the architecture of this implementation and
consider the impact it has from a development point of view. Next we use this
implementation to quantify its merits using a side-by-side comparison of
generated code versus a manual implementation. We consider experimental results
regarding the image size, usage of the network and execution time of the
event-loop. We also take a more analytical approach and look at the handling of
common data and the size of the source code.

\subsection{Architecture}

Figure \ref{fig:architecture} shows the architecture of the implementation.

\begin{figure}[ht]
  \centering
  \scalebox{.85}{
\begin{tikzpicture}[
  node distance=10mm and 20mm,
  ,>=stealth,thick,black!50,text=black,
  every new ->/.style={shorten >=1pt},
]
  \node (dsl)      [user,align=center]                                     {DSL};
  \node (code)     [generated,right=of dsl]                                {code};
  \node (image)    [generated,right=of code]                               {image};

  \node (nodes)    [dom,above right=of dsl,xshift=-20mm]                   {Nodes};
  \node (platform) [platform,below right=of dsl,align=center,xshift=-20mm] {Platform \&\\Language};
  \node (lib)      [native,below right=of code,align=center,xshift=-20mm]  {Software\\Library};

  \draw [->] (dsl)  -> (code)  node [midway,above=2pt] {generation};
  \draw [->] (code) -> (image) node [midway,above=2pt] {compilation};

  \draw [->,dashed] (nodes.south)    -> ++(0,-15mm);
  \draw [->,dashed] (platform.north) -> ++(0,15mm);
  \draw [->,dashed] (lib.north)      -> ++(0,15mm);
\end{tikzpicture}
}
  \caption{Architecture of Prototype Implementation}
  \label{fig:architecture}
\end{figure}

It illustrates the generation of the DSL into compilable code, using the Nodes
domain information, a platform and language description and the native software
library. The applied grayscale indicates the components' platform dependency:
the DSL is platform and language independent and deals with subsystems, such as
the network, in an abstract way. The Nodes domain implements some of these
abstracted subsystems, but still in a platform and language independent way.
The platform and language descriptions define the target platform and language
and convert the abstract constructions in actually supported code. The software
library is standard native code, targeting the platform and language. Finally,
the code and image are artifacts of the automated generation and compilation
process. They aren't a product of manual work and therefore aren't directly
dependent on the platform.

This hierarchy is important from a development effort point of view. The level
of dependency on the targeted platform also dictates the level of effort to
implement changes to that platform. For example: to generate a different
language will require a complete rewrite of the software library and a partial
rewrite of the language description of the generator. Although this is a
reusable, one-time effort, it is important in relation to the scalability of
the framework itself.

\subsection{Implementation}

To evaluate the prototype we chose to use a system based on the Atmel
ATMEGA1284p micro-controller and the Digi XBee S2 ZigBee module. A basic
library wrapped the technical calls to the hardware components in more easy to
use functions.

Relying on basic hardware and software allows us to validate that the generator
is capable of generating code even for the most basic environment available.
Any step up, both in hardware or software, would offer better and higher
abstractions that would make it easier to generate code for that situation. It
shows that the requirements of the generator towards the platform are minimal,
don't rely on advanced frameworks or operating systems and therefore are
applicable in any environment, targeting any platform.

For the evaluation we constructed a small meshed network. In this network, not
all devices have direct access to one another, requiring routing through shared
neighboring devices.

We started from a basic application that measures light intensity. The
application was an includable piece of code, which allowed integration in a
manual implementation and simple merging when generation. On top of this
baseline we added two intrusion detection related algorithms: a
watchdog \cite{mishra2004intrusion}, that allows nodes to validate each other's
presence, and a reputation-based algorithm \cite{ganeriwal2008reputation}, that
checks if a parent node is cooperative and actually forwards messages further
upstream.

\subsection{Experimental Results}

We evaluated three criteria: the image size of the resulting compiled code, the
network usage in number of packets and bytes and the time required to perform
once cycle of the event loop. To do this, we collected these metrics for four
situations: without ID algorithms, with a watchdog, with reputation-tracking
and with both algorithms implemented.

The manual implementation realized both algorithms as standalone modules and
sequentially called them from the base-application's event loop. Given the
\NAME descriptions of the algorithms, the code generator generated all four
cases.

Table \ref{tbl:results} shows the collected data for the manual and the
generated implementation, as well as a comparison in absolute values and
relative.

These results show that adding algorithms manually to the base-application has
a cumulative effect. The impact of both algorithms is simply the sum of the
impact of each algorithm by itself. In case of the execution time of one cycle
of the event loop, the total is even a bit higher.

In case of the generated implementation, this effect is no longer observed:
individual algorithms do have a larger impact by themselves, but the
combination of both algorithms is less than the sum of both.

Most remarkable is the effect on the time of one event loop cycle: both
algorithms by itself add 152\%. Adding a second algorithm only augments the
impact to 188\%. We see the effect of the generic handling of common
functionality, like message parsing, event scheduling,\dots It comes with an
overhead for a single algorithm, but pays for itself when adding more
algorithms.

Table \ref{tbl:results} also compares the two situations by subtracting the
manual case from the generated case, showing the impact of the code generation.
A relative value for this difference, to the manual case, is also included.

The first observation deals with the base case: both the manual and the
generated implementations perform exactly the same. There is no fundamental
structural difference in the overall architecture of the code.

Secondly, we see that the software library that comes with the generated code
adds to the size of the resulting image. But the cost is close to constant and
is lower for both algorithms together. With roughly 3KB of additional overhead,
or 15\% in this simple case, the impact is small. Taking into account the
available memory on the ATMega1284p, this is merely 2.3\% of the available
resources.

Thirdly, when turning to the goal to reduce the network usage, we see that both
the number of packets as well as the amount of bytes sent have dropped by 13
and 7\%.

Finally, focusing on the execution time, both algorithms by themselves add an
equal amount of time to the processing of a single event loop cycle. But when
combined, the event loop is about 7\% faster compared to the manual case.

\subsection{Analytical Results}

Looking at the generated source code is another way to validate the framework.
A simple metric is comparing the amount of code versus the size of the formal
description of IDS algorithms. This can be an indication of the development
effort and complexity. In this experiment, implementing both algorithms
manually results in 637 lines of C code. The corresponding \NAME implementation
results in 156 lines of DSL code.

We can also look at data structures. Sharing data eliminates redundant
information, such as a network address, from the scope of each individual
algorithm. Analysis of the source code with both algorithms implemented, shows
that this reduces the memory footprint for one node entry by 2 bytes, from 27
bytes to 25 bytes. A ZigBee network address indeed consists of 16 bits. So this
results in a reduction of 7.4\%. Using ZigBee's 64 bits unique addresses, this
reduction can easily become 20.5\%.

\section{Related Work}
\label{related}

Section \ref{classification} introduced some ID algorithms
\cite{ganeriwal2008reputation,mishra2004intrusion,krontiris2009cooperative} and
presented a classification
\cite{mishra2004intrusion,ioannis2007towards,alrajeh2013intrusion} based on
their detection approach and interaction style.

This section focuses on related work in the field of domain specific languages
and code generation, both in relation to wireless networks of
resource-constraint devices. These two components are the core of the \NAME
framework and the solution it proposes. We also look at their application in
other disciplines and how this can relate to our context.

\subsection{Domain Specific Languages}

Within the scope of wireless networks of resource-constraint devices, we see
the introduction of DSLs \enquote{to shorten the development cycles
dramatically during deployments, to reduce the dependency on hardware and
software engineers, and to lead to a wider adoption outside the computer
science field} \cite{sadilek2008domain}.

TinyScript \cite{levis2004tinyscript} is an example of a language for wireless
sensor networks. It's a complete programing language that aims to ease the
description of algorithms in general. The Basic-like syntax does enable more
productive development. But this approach doesn't address the underlying
resource problems inherently connected to the context.

SEAL \cite{elsts2013seal} is a novel complete programming language that focuses
on novice programmers. It is a clear fit for \emph{sense-and-send} type of
applications, and doesn't pretend to be the best choice for any type of
application. It does however share the idea of event to overcome the technical
nature of loops.

Shifting the scope to intrusion detection, we learn that DSLs are predominantly
used to describe patterns of events found in network
packets \cite{sekar1999high,roesch1999snort} in a concise way.

BMSL \cite{uppuluri2001experiences} is a DSL that allows the specification of
patterns and actions. Patterns can contain events extracted from log files,
system calls and network packets.

The STAT framework with its STATL language
\cite{eckmann2002statl,vigna2003designing} are not intrusion detection
specific, but through extensions they allow to model attack scenarios using
state machines. Our event-driven approach shares these same ideas.

\subsection{Code Generation}

Code generation, in combination with a DSL, can optimize a development process
by reducing the size of the code base or systematically improving the code
quality.

Embedded systems are a prime example of an environment where code needs to be
of a high quality. Focus on size and performance are a logical choice when
dealing with limited storage and processing power. Compilers implement
different techniques to address this at the lowest level
\cite{marwedel2002code}.

Optimizing for power efficiency however proves to be concurrent to size and
performance optimization, and requires different strategies. Data and memory
optimizations \cite{panda2001data} can result in improvements on a lower level.
But really preserving energy requires higher-level strategies, targeting
architecture, radio handling, efficient software,\dots \cite{naik2001software}.
Most of these are beyond the control of the compiler and require a higher level
of abstraction to address them.

DSLs offer a solution here, but other paradigms are possible. A mixed framework
consisting of a virtual machine (VM) and native code extension
\cite{sadilek2007energy}, combined with energy-aware compilation, seems a
promising approach. By offloading the optimization to a VM and only compiling
parts and not the whole, the solution allows updating at a granular scale, thus
reducing the update cost, while maintaining flexibility and lowering the
development effort. As explained in \ref{positioning}, \NAME focuses on the
reduction of the development effort and update cost, while preserving maximal
flexibility.

Finally, applying code generation to IDS, typically tries to generate native
code to avoid the interpretation of the description of patterns.
I2S \cite{charitakis2003code} is an example that generates code for
SNORT \cite{roesch1999snort} rules. To optimize this highly efficient code, it
often targets a specific platform.

\section{Conclusion and Future Work}
\label{conclusion}

Adequate intrusion detection for IoT devices requires multiple algorithms to
operate at the same time. Recent literature describes different algorithms, but
hardly any are readily available for direct application when implementing an
IDS for IoT devices. Reselection of algorithms, due to changing requirements or
security threats, requires a flexible way to incorporate the different
algorithms. This eliminates manual implementation and optimization.

Wireless resource-constraint devices, such as IoT devices, require that
software respects their restrictions. Adding multiple intrusion detection
algorithm requires fusion of these algorithms on a functional level, to reduce
execution inefficiencies, redundant memory usage and superfluous network access.

\NAME is a framework consisting of a DSL, a source code generator and
accompanying software library. Using the DSL, intrusion detection algorithms
can be formally described in terms of functionality. During source code
generation, functional code fusion interprets the functional meaning of the
algorithms and merges common functionality and shared data. The software
library hosts common functionality, such as network message parsing, sent
message aggregation, event scheduling,\dots This allows the source code
generator to focus on the optimization of the organization of the functionality.

Initial experiments confirm the theoretical positive impact on resource usage.
The platform independent DSL enables generation of source code targeting
different platforms and languages, which allows high degrees of reuse in
heterogenous environments and different configurations.

Future research will focus on the integration of applications with the
generated intrusion detection systems. We see opportunities for applications
who detect intrusion, to actively and dynamically respond and avoid malicious
payloads.

A continued effort is targeting the reduction of platform-dependent parts in
the implementation. This way it is possible to add more platforms and languages
with less effort.

In the longer run, we want to investigate the possibility to extend the
functional code fusion paradigm to other domains beyond intrusion detection.

\section{Acknowledgements}

This research is partially funded by the Interuniversity Attraction Poles
Programme Belgian State, Belgian Science Policy, and by the Research Fund KU
Leuven. This research is partially funded by the EU FP7 project NESSoS\@. We
would like to thank the reviewers for their thoughtful and helpful comments
that enhanced the readability of this paper.

\bibliographystyle{IEEEtran}
\bibliography{literature/referenties}

\end{document}