### 2015-04-17 paper-euspn15.tex
completed text clean-up, squeezed whitespace

% This paper will be submitted to EUSPN15.

\documentclass[3p,times,procedia]{elsarticle}
\flushbottom
\usepackage{ecrc}
\volume{00}
\firstpage{1}
\journalname{Procedia Computer Science}
\runauth{Christophe Van Ginneken, Jef Maerien,
Christophe Huygens, Wouter Joosen, Danny Hughes}
\jid{procs}
\usepackage{amssymb}
\usepackage[figuresright]{rotating}

\input{packages}
\input{squeeze}

\begin{document}

\begin{frontmatter}

\dochead{6th International Conference on Emerging Ubiquitous Systems and
Pervasive Networks, EUSPN-2015 and the 5th International Conference on Current
and Future Trends of Information and Communication Technologies in Healthcare,
ICTH 2015,}

\title{
\FOO: A framework for efficient source code generation\\
of detection algorithms for the Internet of Things
}

\author{Christophe Van Ginneken}
\author{Jef Maerien}
\author{Christophe Huygens}
\author{Wouter Joosen}
\author{Danny Hughes}

\address{
iMinds-DistriNet, KU Leuven, 3001 Leuven, Belgium\\
\{firstname.lastname\}@cs.kuleuven.be
}

\begin{abstract}

The ability to detect patterns in its usage is what makes \emph{things}
\emph{smart}. The more patterns it can detect, the more accurate its
operational results will be. However, supporting an adequate set of pattern
detection algorithms imposes significant overhead on resource-constrained
devices such as those in the Internet of Things. Manual fusion of the
algorithms reduces resource consumption by optimizing resource sharing, yet,
proves to be time-consuming, repetitive and error-prone. To address this
problem we propose \FOO, a framework for the generation of efficient
combinations of detection algorithms. It consists of a domain specific language
and a code generator. The language allows formally describing the intent of
detection algorithms and supports the generator in organizing the source code.
As a case study, we apply \FOO to the generation of an intrusion detection
system for wireless sensor nodes. A side-by-side comparison shows that
\FOO-generated source code reduces message passing overhead, execution time and
memory footprint in comparison to sequential calls into standalone
implementations of the algorithms.

\end{abstract}

\begin{keyword}

Internet of Things, Domain Specific Languages, Source Code Generation, Source
Code Optimization, Intrusion Detection, Context Awareness

\end{keyword}

\end{frontmatter}

\section{Introduction}
\label{introduction}

The Internet of Things (IoT), the mother of all ubiquitous and pervasive
networks, populates our daily lives with millions of internet-connected
\emph{smart things}: from smart watches that assist visually impaired
persons\cite{porzi2013smart} to monitoring systems for waste management in our
\emph{smart cities}\cite{zanella2014internet}. Their applications cannot be
more diverse, however they share one common concern: each of these
\emph{things} is resource-constrained, relying on a battery and operating with
limited processing power and memory.

Although their applications at first seem unrelated, under the hood they share
a common paradigm: pattern recognition. With the introduction of \emph{smart
things}, we want to augment our lives with information that is hard, or at
least time and resource consuming, to obtain otherwise. The ability to detect
patterns in its usage or its applied context, is what makes these \emph{things}
\emph{smart}. A watch that recognizes gestures can perform certain actions on
behalf of its wearer. That same watch can also detect a wet floor and notify
this hazard\cite{porzi2013smart}.

Monitoring human activities is an active research field for more than a
decade\cite{lara2013survey}, but with the advent of the IoT it now moves into
the real world at an increased pace. Equipped with wearables, hosting numerous
sensors, the possibilities seem endless, but so are the number of detection and
recognition algorithms that have to run simultaneously on these small
\emph{things}, to support our augmented reality.

Another prime example of a class of algorithms that inhibit the same properties
is Intrusion Detection (ID). The large number of security related opportunities
for attackers requires an equally sized amount of ID algorithms, all with a
similar structure, still independent in their operation.

% scientific contribution

The contributions of this work are threefold: the first contribution consists
in the formulation of a design pattern for detection/recognition algorithms for
IoT devices. This defines the target algorithms that fit our proposed
framework. The second contribution is a framework called \FOO.\@ It enables the
formal description of algorithms on a functional level and provides a code
generator producing source code for a given platform and configuration. This
way, \FOO also addresses the heterogeneous nature of both IoT and the
algorithms. The third contribution introduces functional code fusion (FCF) as a
source code generation paradigm to address the combination of algorithms. FCF
identifies common data and functions, and organizes code to eliminate redundant
iterations, tests and computations. Results show reduced usage of the wireless
radio, execution time and memory usage.

% structure

The remainder of this paper proceeds as follows: section \ref{problems}
identifies problems that arise when combining multiple algorithms in the
context of resource-constrained devices. Section \ref{pattern} formulates a
pattern common to detection/recognition algorithms in the IoT context. Section
\ref{design} introduces our framework, its design and components. In section
\ref{evaluation} we apply it to a case study related to ID and evaluate the
results. Section \ref{related} explores related work in the field of DSLs and
code generation. Finally, section \ref{conclusion} summarizes our findings,
draws conclusions and identifies topics for future work.

\section{Problems when combining multiple algorithms}
\label{problems}

Simply taking existing implementations of algorithms and calling them in
sequence, will in most cases result in suboptimal code. Three prime examples on
the source code level illustrate this:

\begin{description}[style=unboxed,leftmargin=0cm]

  \item[Loops] by themselves are not a problem, but different loops, that
  perform the same operations in sequence, can cause overhead. An example is
  the handling of messages: each algorithm parses each observed message on the
  network, processing the same byte stream, performing comparable operations to
  find similar patterns. Polling sensors for their value is another example.
  This redundancy results in an execution overhead, putting a strain on the
  power consumption.

  \item[Common data] cross-cuts the different algorithms, that have to store
  the same information over and over again within their own scope. Typical
  examples here include information about neighboring nodes, sensor readings,
  etc.\ This results in a lot of duplicated information in memory, raising the
  hardware requirements.

  \item[The network] allows algorithms to operate on a distributed scale.
  Scattered calls to a message sending function cause multiple accesses to the
  wireless radio, keeping it active and energy consuming more than strictly
  needed.

\end{description}

Two operational aspects are orthogonal to these problems and further amplify
them:

\begin{description}[style=unboxed,leftmargin=0cm]

  \item[Reimplementing] algorithms manually to eliminate these issues is
  not a one-time cost. A release of a new version of an algorithm, or simply a
  reselection of algorithms, due to a changing requirements, can trigger it.
  Also, changing, or simply implementing the algorithms from scratch, holds the
  risk of making mistakes.

  \item[The heterogeneity of IoT devices] results in different software stacks,
  each with for example their own network API\@. This partially explains why
  there are hardly any implementations of algorithms readily available to
  developers: there is no single platform to target for researchers, which
  lowers the opportunity to reuse prior work.

\end{description}

\section{A design pattern for detection/recognition algorithms}
\label{pattern}

The source code level problems that arise when combining multiple algorithms,
as presented in section \ref{problems}, are in fact inherent to the design
pattern that forms the foundation for the type of algorithms we identify in
typical detection/recognition applications in an IoT context. The formulation
of such a design pattern will allow us to define a strategy for the
optimization of the combination of these algorithms. We start by extracting two
common activities: the collection of sensor values and the evaluation of
thresholds. Periodically collecting sensor values requires preparing the
hardware for value extraction, waiting for the value to become available,
reading the bytes from the sensor, interpreting these bytes and recording this
interpretation for further analysis. This can be as simple as reading a single
byte, but can also be as complex as parsing a network message.

A single sensor value hardly ever results in a decision. The combination, often
over time, of these values and the evaluation of thresholds will be key to
decide if a change in the environment requires an action. Updating such
statistics and evaluating the corresponding thresholds should be distinct
operations. This decoupling of detection and decision is the core of the design
pattern for detection algorithms. Listing \ref{alg:pattern} presents the design
pattern in pseudo-code. Applying the pattern in the context of multiple
algorithms, extends its scope and an application pattern such as described in
listing \ref{alg:application} appears.

We see the three aforementioned problems manifest themselves: first, nesting of
loops causes redundant executions of potentially the same sensor readings
inside individual algorithms' $collect\_data$ functions. Second, each
individual algorithm stores the same sensor values in memory. Third, scattered
calls to network functions (in case of distributed algorithms) lead to
suboptimal use of the wireless radio.

When implementing the solution from scratch, it would be possible to identify
these problems and reorganize the code in such a way that it reads and stores
sensor values in memory once or centrally parses network messages. In section
\ref{problems} we identified two operational problems that show that this is
not feasible, or at least suboptimal. We therefore present \FOO, a framework
that enables formally describing such algorithms and fusing common data and
functions in an automated way, to overcome both these problems.

\begin{algorithm}[t]
  \small
  \caption{Detection/Recognition Algorithm Pattern}
  \label{alg:pattern}
  \begin{algorithmic}
    \Require{readings: global variables to hold collected sensor data for the given algorithm}
    \Require{devices: global variable to information about neighbor devices} \Comment{optional in case of distributed algorithms}
    \Function{collect\_data}{}
      \ForEach{$sensor \in required\_sensors$}
        \Let{values[]}{$sensor$}   \Comment{collect sensor data}
        \Let{$reading_{sensor}$}{\Call{interpret}{values[]}}  \Comment{update interpreted sensor info}
      \EndFor
      \State \Call{send}{$devices_x$, ``info''} \Comment{optionally exchange info} \label{alg:id-algo-pattern-send1}
    \EndFunction
    \Function{do\_housekeeping}{}
      \ForEach{$reading \in readings$} \label{alg:id-algo-pattern-loop2} \label{alg:id-algo-pattern-common-data}
        \If{$reading > \dots$} \Comment{validate recorded value}
          \State \dots \Comment{take actions}
          \State \Call{send}{$device_x$, ``info''} \Comment{optionally communicate} \label{alg:id-algo-pattern-send2}
        \EndIf
      \EndFor
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
  \small
  \caption{Application pattern of multiple algorithms}
  \label{alg:application}
  \begin{algorithmic}
    \ForEach{$algorithm \in algorithms$}
      \State \Call{algorithm.collect\_data}{}
    \EndFor
    \ForEach{$algorithm \in algorithms$}        \Comment{at a given interval}
      \State \Call{algorithm.do\_housekeeping}{}
    \EndFor
  \end{algorithmic}
\end{algorithm}

\section{Design}
\label{design}

The previous sections identified loops, common data and the network as three
important code level problems. These problems are inherent to detection
algorithms, but conflict with the target environment, the IoT\@. \FOO provides
a framework to efficiently implement these algorithms and offers a solution for
all identified problems. It consists of a 1) DSL, 2) source code generator and
3) software library.

The DSL enables researchers to formally describe algorithms, without focusing
on a specific language or platform. The code generator can avoid the identified
problems concerning loops, common data and network access, because the formal
description provides the underlying intent, not the technical implementation.
It applies Functional Code Fusion (FCF) to identify hotspots for optimization.
A key component of FCF are \emph{functional domains} that harbor common
functionality.

We will first position \FOO as a development strategy and motivate its design.
Second, we will look at the principal design components of the DSL.\@ Third, we
discuss the structure of the code generator and introduce its software library.

\subsection{\FOO as a development strategy}
\label{positioning}

The introduction of a code generator in the development process changes the way
of working: the source code is no longer the driving force, controlled by
developers. It is literally an intermediate representation between code
generator and compiler. A code generator constructs this representation using
three components, as shown in figure \ref{fig:design}: 1) the description of
the algorithm in the DSL, 2) information about the functional domains, and 3)
information about the platform and target language. At the level of the
compiler, a software library offloads specialized implementations and allows
custom extensions. The applied grayscale indicates the components' platform
dependency. This hierarchy is important from a development effort point of
view. The level of dependency on the target platform dictates the level of
effort to implement changes. This is important in relation to the scalability
of the framework itself.

\begin{figure}[t]
  \centering
  \scalebox{0.70}{
\begin{tikzpicture}[
  node distance=9mm,
  ,>=stealth,very thick,black!50,text=black,
  every new ->/.style={shorten >=1pt}
]
  \node (dsl)     [user]                                      {DSL};
  \node (cg)      [representation,right=of dsl,align=center]  {Code\\Generator};
  \node (code)    [generated,right=of cg]                     {code};
  \node (comp)    [representation,right=of code]              {Compiler};
  \node (image)   [generated,right=of comp]                   {image};

  \node (domains) [dom,above=of cg, align=center]             {Functional\\Domains};
  \node (plat)    [platform,below=of cg,align=center]         {Platform \&\\Language};          
  \node (lib)     [native,below=of comp,align=center]         {Software\\Library};

  \draw [->] (dsl)     -> (cg);
  \draw [->] (domains) -> (cg);
  \draw [->] (plat)    -> (cg);
  \draw [->] (cg)      -> (code);
  \draw [->] (code)    -> (comp);
  \draw [->] (lib)     -> (comp);
  \draw [->] (comp)    -> (image);
\end{tikzpicture}
}
\caption{The high level design of \FOO, showing the resources, processes and
artifacts. A grayscale background represents the level of platform dependency,
with darker shades indicating a higher dependency. Both the source code and the
installable image are artifacts.}
\label{fig:design}
\end{figure}

\subsection{\FOO as a domain specific language}
\label{dsl-design}

It's important to remember that a DSL is merely a \emph{thin veneer over a
(semantic) model} \cite{fowler2010domain}. Its sole purpose should be
populating the model. \FOO's DSL and model focus on avoiding constructions that
result in problems concerning loops, common data and the use of the network. To
do this, \FOO uses an event-based approach to defining functionality.
Describing algorithms based on events removes the technical interpretations and
allows FCF to identify the intent of algorithms and avoid the identified source
code problems.

\subsubsection{Loops}

Loops are a major contributor to the problem of algorithm fusibility. Loops in
program code are technical constructs and in fact often hide the intended
functionality. The example of parsing network messages illustrates this. Each
algorithm deals with parsing in the same way: loop over the bytes in the
payload and look for a pattern that identifies what the algorithm is looking
for. This is in fact a technical implementation choice. The underlying
functional intent is to react when such a pattern is present in a message. The
loop has \emph{activated} this functionality, but the real goal is to
\emph{passively} react on an event, not \emph{actively} look for it. Event
handling therefore can transform constructions with loops to their functional
meaning, allowing identification and fusion of common functionality.

\subsubsection{Common data}

Different detection algorithms typically deal with the same classes of
information, be it sensor data or network messages. Functional domains
consolidate these and store and manage this data in a unified way. \FOO
provides the concept of such domains as plug-ins to the DSL, extending it with
domain specific storage and logic. For example in the case of distributed
algorithms a \emph{Nodes} domain will manage information about neighbor nodes,
exposing them as global objects to the algorithms, relieving them from managing
these themselves. Functional domains are also accessible as collections,
allowing execution of functions in their scope, again eliminating the need for
explicit loops.

\subsubsection{Unified messaging}
\label{dsl-unified-msg}

Access to the network, scattered around an algorithm, causes the wireless radio
to be more active than needed. Delaying the sending of messages and grouping
them can lower this overhead. Sending and receiving of network messages is
tightly integrated with the previously mentioned concepts of event handling and
for example the \emph{Nodes} domain. An underlying software library can handle
all interaction with the network, triggering functions by means of events when
parsing messages and exposing an API to send messages. This approach gives the
framework complete control over communication and allows for optimized
marshaling and grouping of messages.

\subsection{\FOO as a code generator}

To convert the DSL-based ID algorithms into production-ready code, \FOO applies
code generation techniques. Although different approaches are applicable, like
template-based text expansion, XSLT transformations or CodeDom
\cite{dollard2004code}, we believe that the required transformations demand a
richer model. A semantic model \cite{fowler2010domain}, tied closely to the DSL
itself, can fulfill this role. Semantic transformations evolve the model into a
more syntax-oriented model, comparable to the idea of CodeDom. Syntactic
transformations can further evolve the model into source code. Figure
\ref{fig:code-generation} shows this process.

\begin{figure}[t]
  \centering
  \scalebox{.70}{
\begin{tikzpicture}[
  node distance=7mm,
  ,>=stealth,thick,black!50,text=black,
  every new ->/.style={shorten >=1pt}
]
  \node (dsl)  [resource]                                 {DSL};
  \node (sm)   [representation,right=of dsl,align=center] {Semantic\\Model};
  \node (cm)   [representation,right=of sm,align=center]  {Syntactic\\Model};
  \node (code) [resource,right=of cm]                     {Source Code};

  \draw [->] (dsl) -> (sm)   node [midway,above=15pt]              {parse};
  \draw [->] (sm)  -> (cm)   node [midway,below=10pt,align=center] {semantic\\transform}
                             node [midway,above=15pt,align=center] {FCF};
  \draw [->] (cm)  -> (code) node [midway,above=15pt]              {emit};

  \draw [rounded corners,->]
        ($ (sm.east) - (5mm,5mm) $)
        -- ++(0,-.5)
        -| ($ (sm.west) + (5mm,-5mm) $)
           node [near start, below=5pt] {type inference};

  \draw [rounded corners,->]
        ($ (cm.east) - (5mm,5mm) $)
        -- ++(0,-.5)
        -| ($ (cm.west) + (5mm,-5mm) $)
           node [near start, below=1pt,align=center] {syntactic\\transform};

\end{tikzpicture}
}
\caption{\FOO's code generation process}
\label{fig:code-generation}
\end{figure}

\subsubsection{Semantic model}

The semantic model is the core of the generator. It offers a way to describe
algorithms so that semantic reinterpretation through FCF is possible. This
reorganization of the functionality of different algorithms allows for more
optimal execution and memory usage. Hence the name \FOO of the framework:
functionality organization optimization. Figure \ref{fig:meta-model} shows the
central part of the semantic model as implemented for \FOO.\@ The concept driving
this model is the structure starting at the \emph{Model} up to the
\emph{Execution Strategy}. This structure encompasses the \emph{Function}s and
underlying \emph{Statement}s and allows for interpretation and fusion of the
algorithms.

\begin{figure}[b]
  \centering
  \scalebox{.70}{
\begin{tikzpicture}[
  edge from parent fork down,
  sibling distance=5mm, level distance=20mm, node distance=6mm
  ]
  \node (model) [class,fill=black!10!white] {Model}
      child {node (module) [class,fill=black!10!white] {Module} edge from parent [composite]
        child {node (strategy) [class,align=center,xshift=-10.5mm,fill=black!10!white] {Execution\\Strategy} edge from parent [composite]
          child {node(every) [class,xshift=-9mm,fill=black!10!white] {Interval} edge from parent [subclass]}
          child {node (when) [class,xshift=9mm,fill=black!10!white] {Event} edge from parent [subclass]}
        }
        child {node (function) [class,xshift=10.5mm] {Function} edge from parent [composite]}
      };
  \node (constant)  [class,right=of module]                {Constant};
  \node (scope)     [class,left=of strategy]               {Scope};
  \node (statement) [class,right=of function]              {Statement};
  \node (domain)    [class,left=of module,xshift=-13mm]    {Functional\\Domain};

  \draw [aggregate]  (module)   -- (domain);
  \draw [composite]  (module)   -- (constant);
  \draw [dependency] (strategy) -- (scope);
  \draw [composite]  (function) -- (statement);

  \draw [composite]  (domain)   -- (scope);
  \draw [dependency] (strategy) -- (function);
\end{tikzpicture}
}
\caption{Partial Semantic Model for \FOO.\@ Shaded classes indicate the
\emph{event/scheduling} construction around functions.}
\label{fig:meta-model}
\end{figure}

\subsubsection{Functional code fusion (FCF)}

Thanks to the functional description of events and the effects triggered by
them, the code generator is able to identify the same actions across different
algorithms. This way it is possible to extract the parsing of incoming messages
and group actions on the same sets of data or scheduled function executions.
FCF typically looks for patterns in the semantic model and performs model
transformations to change these patterns to a more optimized organization. It
first performs semantic transformations on the semantic model itself. A second
phase transforms the semantic model into a syntactic model, encoding the
semantics into technical syntax constructs. One example of a semantic
transformation is type inference, which supports the platform independence of
the framework.

\subsubsection{Syntactic model}

The syntactic model is like an abstract syntax tree for a virtual, expressive
language. It contains constructs and ideas found in different languages and
programming paradigms. No single existing programming language supports all
features as is. Syntactic transformations therefore translate the unsupported
constructs into comparable constructs supported by the targeted platform and
language. Depending on the target platform, some constructions aren't
transformed into actually corresponding code. In selected cases, the generator
uses a software library. For example when it needs to parse incoming messages
or schedule tasks. During the generation process, it looks for patterns that
indicate any of these actions and transforms them into calls to the \FOO
software library.

\subsection{\FOO as a software library}
\label{software-lib-design}

A software library allows offloading of common implementation aspects, thus
avoiding small discrepancies in code and suboptimal use of shared resources.
Previously we illustrated this approach with a case for unified messaging:
parsing of incoming messages and grouping of outgoing messages. More common
functionality isn't generated verbatim either. A standard software library
provides functionality for the execution of functions based on a schedule or
the implementation of higher level conceptual functions. Optimization of these
type of algorithms is not within the scope of \FOO.\@ This further gives
freedom of choice to the users of the framework to select appropriate
implementations for this standard functionality.

\section{
Implementation and evaluation:
a case study on intrusion detection in wireless sensor networks
}
\label{evaluation}

In wired networks, firewalls focus on the outer perimeter of the network,
filtering unwanted packets. Still, attacks on services can pass unnoticed. This
is where an Intrusion Detection System (IDS) comes into play: monitoring all
traffic that passes through the firewall, looking for patterns of malicious
activity\cite{denning1987intrusion}. In meshed networks it is not possible for
a single point in the network to oversee all traffic
\cite{mishra2004intrusion}. So every device has to implement its own IDS.\@

Attackers have access to a large set of possible attack vectors
\cite{aschenbruck2012security}, ranging from the physical layer up to the
application layer. Each layer presents different opportunities for launching an
attack. For each of these attacks an algorithm needs to look for patterns or
anomalies. An IDS therefore requires a lot of algorithms to achieve adequate
coverage.

\subsection{Implementation}

To evaluate the prototype we used a system based on the Atmel ATMEGA1284p
micro-controller and the Digi XBee S2 ZigBee module, running no specialized
operating system. Relying on basic hardware and software allows us to validate
that the generator is capable of generating code even for the most basic
environment. For the evaluation we constructed a small meshed network. In this
network, not all devices have direct access to one another, requiring routing
through shared neighboring devices.

We started from a basic application that measures light intensity. The
application was an includable piece of code, allowing integration in both a
manual implementation and generated code base. On top of this baseline we added
two intrusion detection related algorithms: a watchdog
\cite{mishra2004intrusion}, that allows nodes to validate each other's
presence, and a reputation-based algorithm \cite{ganeriwal2008reputation}, that
checks if a parent node is cooperative and actually forwards messages.

\subsection{Analytical results}

Looking at the generated source code, a simple metric is comparing the amount
of code versus the size of the formal description of ID algorithms.
Implementing both algorithms manually results in 637 lines of C code. The
corresponding \FOO implementation requires only 156 lines of DSL code. Looking
at data structures, we see that sharing data eliminates redundant information,
such as a network address. This reduces the memory footprint for one node entry
by 2 bytes, from 27 bytes to 25 bytes. In our implementation we used the ZigBee
network address, which consists of 16 bits. This results in a reduction of
7.4\% of dynamic memory per known device. Using ZigBee's 64 bits unique
addresses, this reduction can easily become 20.5\%.

\subsection{Experimental results}

We also instrumented the code and evaluated four criteria: network usage
(number of packets and bytes), the time required to perform once cycle of the
event loop and the resulting image size. We collected these metrics for four
situations: without ID algorithms, with a watchdog, with reputation-tracking
and with both algorithms. The manual implementation realized both algorithms as
standalone modules and sequentially called them from the base-application's
event loop. Given the \FOO descriptions of the algorithms, the code generator
generated all four cases. Table \ref{tbl:results} shows the collected data.

\begin{table}[ht]
  \caption{Experimental results and side-by-side comparison.}
  \label{tbl:results}
\begin{tabular*}{\hsize}{@{\extracolsep{\fill}}llrrrrrrr@{}}
\toprule
              &            & base  & \multicolumn{2}{c}{watchdog} & \multicolumn{2}{c}{reputation} & \multicolumn{2}{c}{both} \\
\midrule
\multirow{2}{*}{\pbox{\textwidth}{total packets sent\\in a window of 90s}}
              & manual     &    20 &    51 & +155\% &    32 &  +60\% &    63 & +215\%\\
              & generated  &    20 &    49 & +145\% &    32 &  +60\% &    55 & +175\%\\
\cmidrule{2-9}
              & difference &     0 &    -2 &   -4\% &     0 &   0\% &    -8 &   -13\%\\
\midrule
\multirow{2}{*}{\pbox{\textwidth}{total bytes sent\\in a window of 90s}}
              & manual     &   476 &  1933 & +306\% &   860 &  +81\% &  2317 & +387\%\\
              & generated  &   476 &  1897 & +299\% &   884 &  +86\% &  2161 & +354\%\\
\cmidrule{2-9}
              & difference &     0 &   -36 &   -2\% &    24 &   +3\% &  -156 &   -7\%\\
\midrule
\multirow{2}{*}{\pbox{\textwidth}{event loop\\average duration ($\mu$s)}}
              & manual     &    48 &    94 &  +96\% &    88 &  +83\% &   149 & +210\%\\
              & generated  &    48 &   121 & +152\% &   121 & +152\% &   138 & +188\%\\
\cmidrule{2-9}
              & difference &     0 &    27 &  +29\% &    33 &  +38\% &   -11 &   -7\%\\
\midrule
image
size (bytes)  & manual     & 10466 & 15530 &  +48\% & 13306 &  +27\% & 18334 &  +75\%\\
              & generated  & 10466 & 18352 &  +75\% & 16376 &  +56\% & 20998 & +101\%\\
\cmidrule{2-9}
              & difference &     0 &  2822 &  +18\% &  3070 &  +23\% &  2664 &  +15\%\\
\bottomrule
  \end{tabular*}
\end{table}

It is clear that adding algorithms manually to the base-application has a
cumulative effect. In case of the generated implementation, this effect is no
longer observed: individual algorithms do have a larger impact by themselves,
but the combination of both algorithms is less than the sum of both. Most
remarkable is the effect on the time of one event loop cycle: both algorithms
by itself add 152\%. Adding a second algorithm only augments the impact to
188\%. We see the effect of the generic handling of common functionality. It
comes with an overhead for a single algorithm, but pays for itself when adding
more algorithms.

Table \ref{tbl:results} also compares the manual versus de generated versions,
showing the impact of the code generation: both the number of packets as well
as the amount of bytes sent have dropped by 13\% and 7\%. In case of the
execution time, both algorithms by themselves add an equal amount of time to a
single event loop cycle. When combined, the event loop is about 7\% faster
compared to the manual case. Finally, we see that the software library adds to
the size of the resulting image. However, the cost is close to constant and is
lower for both algorithms together. With roughly 3KB of additional overhead, or
15\% in this simple case, the impact is small. Taking into account the
available memory on the ATMega1284p, this is only 2.3\% of the available
resources.

\section{Related work}
\label{related}

This section focuses on related work in the field of domain specific languages
and code generation, in relation to wireless networks of resource-constrained
devices. These two components are the core of the \FOO framework.

Within the scope of wireless networks of resource-constrained devices, we see
the introduction of DSLs \enquote{to shorten the development cycles
dramatically during deployments, to reduce the dependency on hardware and
software engineers, and to lead to a wider adoption outside the computer
science field} \cite{sadilek2008domain}. TinyScript \cite{levis2004tinyscript}
is an example of a language for wireless sensor networks. It's a complete
programing language that aims to ease the description of algorithms in general.
The Basic-like syntax does enable more productive development. Still, this
approach doesn't address the underlying resource problems inherently connected
to the context. SEAL \cite{elsts2013seal} is a novel complete programming
language that focuses on novice programmers. It is a clear fit for
\emph{sense-and-send} type of applications. It does however share the idea of
using events to overcome the technical nature of loops. In the context of
intrusion detection, we learn that DSLs are predominantly used to describe
patterns of events found in network packets
\cite{sekar1999high,roesch1999snort} in a concise way. The STAT framework with
its STATL language \cite{eckmann2002statl,vigna2003designing} allows modeling
scenarios using state machines. Our event-driven approach shares this idea.
Although not a real DSL, nesC \cite{gay2003nesc} is a general purpose language
that shares some typical concepts with \FOO, such as the high degree of
analysis during code generation to optimize execution. The biggest difference
is that \FOO starts from the intended functionality and can optimize certain
constructions that can't be interpreted from more technical source code.

Code generation, in combination with a DSL, can optimize a development process
by reducing the size of the code base or systematically improving the code
quality. Embedded systems are a prime example of an environment where code
needs to be of a high quality. Focus on size and performance are a logical
choice when dealing with limited storage and processing power. Compilers
implement different techniques to address this at the lowest level
\cite{marwedel2002code}. Optimizing for power efficiency however proves to be
concurrent to size and performance optimization, and requires different
strategies. Data and memory optimizations \cite{panda2001data} can result in
improvements on a lower level. To really preserve energy one requires
higher-level strategies, targeting architecture, radio handling, efficient
software,\dots \cite{naik2001software}. Most of these are beyond the control of
the compiler and require a higher level of abstraction to address them. DSLs
offer a solution here, but other paradigms are possible. A mixed framework
consisting of a virtual machine (VM) and native code extension
\cite{sadilek2007energy}, combined with energy-aware compilation, seems a valid
approach. By offloading the optimization to a VM and only compiling parts and
not the whole, the solution allows updating at a granular scale, thus reducing
the update cost, while maintaining flexibility and lowering the development
effort. In this terminology, \FOO focuses on the reduction of the development
effort and update cost, and maximizes flexibility.

\section{Conclusion and future work}
\label{conclusion}

Many IoT applications typically are constructed of similar algorithms that
adhere to a pattern of collecting sensor data and periodically checking
recorded values for patterns that require action. Fusing these algorithms is
hard but needed to deal with the resource-constrained devices they are
targeting. Frequent reselection, e.g. due to changing requirements or changing
security threats, requires a flexible way to incorporate these algorithms in
subsequent iterations of the software.

\FOO is a framework consisting of a DSL, a source code generator and software
library. The DSL allows formally describing algorithms, enabling the code
generator to apply functional code fusion to interpret the intent of algorithms
and merge common functionality \& shared data. The software library hosts
common functionality, such as sensor access, message parsing and event
scheduling. The platform independent DSL enables generation of source code
targeting different platforms and languages, which allows high degrees of reuse
with heterogenous devices.

Initial experiments confirm the theoretically positive impact on resource
usage: grouping of messages lowers the number of packets by 13\%, while the
total number of bytes sent over the network is reduced by 7\%. The execution
time is improved by 7\% and memory usage can be reduced even by 20\%.

Future research will focus on the integration of applications with the
generated functionality. We see opportunities for applications to adopt the
event-driven approach and hope to expand the capabilities of \FOO to this
level. A continued effort is targeting the reduction of platform-dependent
parts in the implementation. This way it is possible to add more platforms and
languages with less effort. In the longer run, we want to investigate the
possibility to extend the functional code fusion paradigm to more domains.

\section{Acknowledgements}

This research is partially funded by the Research Fund KU Leuven and partially
by the EU FP7 project NESSoS\@.

\bibliographystyle{elsarticle-num}
\bibliography{literature/referenties}

\end{document}